---
title: "DHS Datasets"
author: "jp"
date: "10/5/2016"
output:
  html_notebook: default
  pdf_document: default
---

```{r, include= FALSE}

library(knitr)
library(scales, quietly = TRUE)
library(data.table)
library(dtplyr)
library(tidyverse, quietly = TRUE)
library(countrycode)

library(lme4)
library(rethinking)

library(GGally)

opts_chunk$set( echo = TRUE, cache = TRUE )

# location of DHS folder.  include DHS files as: paste0( dhs, <dhs file here>)
dhs = "../DHS/"
```

# Survey datasets

## Select country/surveys to include 

```{r}

source( paste0( dhs, "file_list.r") )

subsahara = countrycode_data %>%
  filter( region %in% c("Middle Africa", "Western Africa", "Eastern Africa", "Southern Africa")) %>% 
  select( country.name, iso3c ) %>%
  filter( !country.name %in% "Lesotho") 


files = files %>% 
  mutate( country.name = countrycode(country, "country.name", "country.name")) %>%
  filter( survey!="", year >= 2000 , country.name %in% subsahara$country.name)

# View(files)

# not_needed = c('Bangladesh', 'Pakistan', 'Philippines', 'Comoros')

survey_list = files %>% count(country, survey, year) %>% select(-n)
# View(survey_list)

nsurveys = nrow(survey_list)
ncountries = length( unique( files$country ) )

cat( "There are", nsurveys, "surveys since 2000 from",  ncountries ,"countries in Subsaharan Africa.")

# kable( survey_list )
```

## Surveys with parasitemia biomarkers

```{r}
dd <-readRDS( paste0( dhs, "dictionaryNew.rds")  )  # about 6 sec

parasitemia_variables = c("hml32", "hml35")

parasitemia_surveys = dd %>% 
  filter( survey != "", year >= 2000 , iso3c %in% subsahara$iso3c ) %>%
  filter( Item_Name %in% toupper( parasitemia_variables) ,
          value %in% 0:1 ) %>% # comment this line out to get full list of variables
  filter( !grepl( "NA", Item_Label, ignore.case = FALSE, fixed = TRUE)  ) %>%
        select( Item_Name, country, survey, value, year) %>%
      count( country, survey, year, Item_Name) %>%
    mutate( 
        has = ifelse( n > 0 , "Y", "N" ) 
        ) %>%
    select( -n) %>%
  spread( Item_Name ,  has, fill = "N" )

 kable(parasitemia_surveys, row.names = TRUE)  
  
```

# Household member dataset

## Select **individual** variables of interest for bednet study and check to see which survey file has the values

```{r}

individual_vars  = c( 
        'hml32', 'hml35', # RDT and micrscopy
        'hml1', "hml4", "hml5", "hml6", "hml7", "hml8",  # bednets
        'hml19', 'hml20', 'hml12', # slept under llin
        'hv102', # de jure member of household
        'hv103', # de facto memeber of household (slept there last night)
        'hml16', # corrected age in years
        "hv253", # dwelling sprayed (some multiple responses)
        "hv001", "hv002", 'HV003', 'HHID', 'HVIDX',
        "HV008", # date of interview
        "HV270", # wealth index
        "latitude", "longitude"
       )
  
variables <- unique( toupper( individual_vars) )
variables = variables[ order(variables) ]

# translation table of dhs file codes to file names
filecodes = data_frame( abrev  = c("kr", "br", "ir", "pr", "hr", "hr" ), 
                            file = c("Children's Recode","Supplemental Births Recode",
                                     "Individual Recode",
                                     "Household Member Recode", "Household Recode", 
                                     "Supplemental Household Recode")) 

 
# list variables and the files where each is found
survey_files = dd %>% semi_join( parasitemia_surveys ) %>%
    filter( Item_Name %in% variables ) %>%
    filter( !grepl( "NA", Item_Label, ignore.case = FALSE, fixed = TRUE) ,
            !grepl( "Presence", Item_Label, ignore.case = FALSE, fixed = TRUE)
            ) %>%
    left_join( filecodes , by = "file", copy = TRUE) %>%
    mutate( 
        abrev = ifelse( is.na(abrev) , tolower( substr(file, 3, 4) ) , abrev)
        ) %>%
    left_join( filecodes , by = "abrev", copy = TRUE) %>%
    rename( file = file.y ) %>%
    count( Item_Name, Item_Label, country, survey, year, file ) %>%
    mutate( has = ifelse( n > 0, "Y", NA )) %>% 
    ungroup() %>%
    count( Item_Name, file ) %>%
    spread( file, nn, fill = "N" )

# View(dataset_surveys)

kable(survey_files)
```

## variable definitions...

Select the first available definition from dictionary.  Note that some variables may be defined differently for some countries, which will not be apparent here. 

```{r}
variable_defn = dd %>% semi_join( parasitemia_surveys, by = c("country", "survey", "year") ) %>%
    filter( Item_Name %in% variables ) %>%
    filter( !grepl( "NA", Item_Label, ignore.case = FALSE, fixed = TRUE) ,
            !grepl( "Presence", Item_Label, ignore.case = FALSE, fixed = TRUE)
            ) %>%
    left_join( filecodes , by = "file", copy = TRUE) %>%
    mutate( 
        abrev = ifelse( is.na(abrev) , tolower( substr(file, 3, 4) ) , abrev)
        ) %>%
    left_join( filecodes , by = "abrev", copy = TRUE) %>%
    rename( file = file.y ) %>%
    filter( grepl( "member", file, ignore.case= T)) %>%
    count( Item_Name, Item_Label) %>%
    select( -n) %>%
    arrange( Item_Label ) %>%
    as.data.frame()

variable_defn
```


## create household member dataset. run one time after updating variables or adding surveys

```{r, eval=FALSE, include = FALSE}
source( paste0( dhs, "getSurveyGIS.R") )
source( "load_survey_file.R") 

nsurveys = nrow( parasitemia_surveys )

parasitemia_surveys = arrange( parasitemia_surveys, desc(year) )
index = row.names(parasitemia_surveys)

p <- progress_estimated(nsurveys)

x <- list()

translate = FALSE  # optional translation of values to label.  
  # NB: Does not work well for all variables

for (i in seq_along( index )  ){
  
   p$pause(0.1)$tick()$print()
  
  .country = parasitemia_surveys[i, ]$country
  .iso3c = countrycode( parasitemia_surveys[i, ]$country, "country.name", "iso3c")
  .survey = parasitemia_surveys[i, ]$survey
  .year = parasitemia_surveys[i, ]$year
  .file = "Household Member Recode"
  .survey_year = paste(.survey, .year)

  print( paste(.country, .survey, .year, .file) )
  
  x[[i]] <- try( silent = TRUE,
    load_survey_file( 
                        # printout = TRUE ,  # un comment when testing
                        vars = variables, 
                        .country = .country, 
                        .survey = .survey,
                        .year = .year,
                        .file = .file ,
                      geo = TRUE)
  )
  
      
    if ( class(x[[i]]) == "try-error" )  next()
     

  x[[i]]$country = rep( .country, nrow(x[[i]]) )
  x[[i]]$iso3c = rep( .iso3c, nrow(x[[i]]) )
  x[[i]]$year = rep( .year, nrow(x[[i]]) )
  x[[i]]$survey = rep( .survey, nrow(x[[i]]) )

  # TODO: optional- relate dictionary to survey                    )

  # # TODO: only for columns with >4 values (others easier to keep as numeric)
  # # Deriving this table should happen once, before loop
  # v = dd %>% filter( tolower(Item_Name) %in% variables) %>% group_by( Item_Name, label ) %>% 
  #   summarise( n = n()) %>% group_by( Item_Name) %>% count(Item_Name) %>% filter(nn>4) 

  if (translate){
    
      # refine dictionary for this country-survey 
      dict = dd %>% filter_( ~country == .country , ~survey == .survey, ~year == .year,
                        ~Item_Name %in% toupper(names(has_vars))
      )
     
    y = x[[i]]
  
    for (col in 1:length(colnames(y))){
      column = tolower( colnames(y)[col] )
      
      dict_var = dict %>% 
        mutate( item_name = tolower(Item_Name)) %>%
        filter_( ~item_name == column ) %>% # , ~file == "Household Member Recode"
        count( value, label) %>% select(-n)
      
      old_value = y[, col]
      new_value = dict_var[ match( old_value, dict_var$value ), ]$label
      y[, col] = ifelse( is.na( new_value), old_value, new_value)
    }
      x[[i]] = y
}

  # if (is.null(X) ){ X <- x; next()}
  # X = rbindlist( list(X, x), use.names = TRUE, fill = TRUE)
  
}
  
# X = dplyr:: bind_rows( x )  # will throw error is column types dont match 
X = rbindlist(x, fill = TRUE) 

# De Facto members only

defacto = X %>% count(hv103)

X = X %>%  filter( hv103 == 1 ) # de facto members only (i.e. slept in house previous night)

 saveRDS(X, "household_member_bednet.rds")
 # comma(object.size(X))

```

# create individual file by filtering household member file to those children with parasitemia result

```{r, eval= FALSE }
hm = readRDS( "household_member_bednet.rds"  ) %>% as_tibble()

 # These countries were supposed to have results, but there are none...
hm %>% 
    mutate(
        RDT = hml32 %in% 0:1 ,
        slide = hml35 %in% 0:1 
    ) %>%
    count( country, survey, year, RDT, slide ) %>% ungroup %>%
    count( country, survey, year ) %>% filter(nn == 1)
    
    
# filter to those with parsitemia result

has.test.result = hm$hml32 %in% 0:1 | hm$hml35 %in% 0:1
i = hm[ has.test.result ,  ]

cat( "There are", nrow(i), "children with a either a positve or negative parsitemia result from",
     count(i, country, survey, year) %>% nrow, "surveys in ",
     count(i, country) %>% nrow, "countries" 
     )
    
    
saveRDS(i, "individual_bednet.rds" )

```

# Create household dataset by aggregating household member file 

net variables:
- hml1: number of nets
- hml4: months ago obtained net
- hml5: was net treated with insecticide when obtained (binary)
- hml6: Mosquito net treated with insecticide (6 options)
- hml7: brand of bednet
- hml8: bednet treated since obtaining

```{r, eval=FALSE}

hm = readRDS( "household_member_bednet.rds"  ) %>% as_tibble()

members = hm %>% 

    count( country, iso3c, survey, year, hv001, hv002) %>%
    rename( household_num = n ) %>%
    ungroup

summary( members$household_num )
quantile( members$household_num , probs = seq(0,1,.1))

# calculate avg household
# h %>% group_by(country, iso3c, survey, year) %>% summarise( 
#     avg_household = mean(n) 
#     )

nets = hm %>% 
    mutate( nets_num_household = ifelse( hml1 %in% 98:99, NA, hml1)) %>%  # 98=don't know; 99 = missing
    group_by( country, iso3c, survey, year, hv001, hv002) %>%
    summarise( nets_num_household = max(nets_num_household, na.rm = TRUE) ) %>%
    ungroup

summary(nets$nets_num_household)  # NB: there are 6 houeholds with NA


h = inner_join( members, nets, 
                by = c("country", "iso3c", "survey", "year", "hv001", "hv002")) %>%
    mutate(
        nets_person = nets_num_household / household_num
    )

summary( h$household_num )
quantile( h$household_num , probs = seq(0,1,.1))

summary( h$nets_num_household )
quantile( h$nets_num_household , probs = seq(0,1,.1))

percentiles = sapply( h[, c("household_num", "nets_num_household")], 
        function(x) quantile(x, probs = seq(0,1,.01), na.rm = TRUE) ) %>%
    as.data.frame() %>% 
    mutate( percentile = factor(row.names(.) , levels=row.names(.) ) ) %>%
    gather(var, value, -percentile)
    

ggplot( percentiles, aes( x = value, y = percentile, group = var)) + 
    geom_line() + scale_y_discrete( breaks = percentiles$percentile[1 +seq(0,100,10)] ) + 
    facet_wrap(~var, scale = 'free')


cat("there are", nrow(h), "households")

saveRDS( h, file = "household_bednet.rds" )
```


```{r}

h = readRDS("household_bednet.rds" )

summary( h$nets_num_household )

summary( h$household_num)

summary( h$nets_person)

# pairs( h %>% select( nets_num_household, household_num, nets_person) )


ggpairs(h, columns=c("nets_num_household", "household_num", "nets_person"))


```


# Load datasets and see which surveys are included


```{r }

i = readRDS(  "individual_bednet.rds" ) %>% as_tibble()
h = readRDS(  "household_bednet.rds" ) %>% as_tibble()

# v001 & hv001 - cluster number
# v002 & hv002 - household number
# b16 & hvidx - line number of child in household (for linking moms and kids)

#join childrens to household to get # persons per household, then join that with household member file.

ih = inner_join( i, h, by = c( "country" = "country",
                               "iso3c"="iso3c",
                              "year" = "year",
                              "survey" = "survey",
                              "hv001" = "hv001",
                              "hv002" = "hv002")
)

data_ncountries = length( unique(ih$country))
data_nsurveys = nrow( count( ih, country, survey, year)  )
nvars = ncol(ih)
rows = nrow(ih)

cat( "The combined individual/household file includes", data_nsurveys, "surveys from", data_ncountries, "countries with", nvars, "columns, ", rows, "rows and is " , comma(object.size(ih)[1]), "bytes")

# list surveys included
kable( ih %>% count( country, survey, year ) , row.names = TRUE)
```

# Individual risk-ratio for bednets (fit0)

with several methods for calculating effect

```{r}



## Slept under net
# filter( dd, Item_Name == "HML12") %>% count(value, label)
slept = integer( nrow(i))
slept[ i$hml12 %in% 1:5] = 1

ts = table(slept, useNA = 'always') 
ts

p = function(n0, yes){ yes/(n0 + yes)}

p(ts[1], ts[2])



## is parasitemic by rdt or slide
hml3235 = integer( nrow(i))
hml3235[ i$hml32 == 1 | i$hml35 == 1 ] = 1

tp = table(hml3235, useNA = 'always')
tp
p(tp[1], tp[2])


# dataset combining variables
d = cbind( hml3235, slept) %>%  as.data.frame()

# tally
count(d, hml3235, slept) 



```


## epitools package

estimates crude RR 

```{r}
library(epitools)
r = riskratio(d$slept, d$hml3235)  
# str(r)
r$measure[2,]

```

## GLM logistic (OR) 

GLM gives estimate of overall risk, and risk ratio

```{r}
library(base)
# logistic regression for rr
glm.logistic = glm( hml3235 ~ slept , data = d, family = binomial(link = "logit"))

### Risk (intercept) and odds ratios 
exp(coef(glm.logistic))

```


### classical log-binomial model (RR)

```{r}
# dataset combining variables
d = cbind( hml3235, slept) %>%  as.data.frame()


# log-binomial regression for rr
glm.log.binomial = glm( hml3235 ~ slept  , data = d, family = binomial(link = "log"))

### Risk (intercept) and risk ratios (under poisson assumption)
exp(coef(glm.log.binomial))

```


## Bayesian, maximum a postiori (MAP) rethinking package

estimates both risk (intercept)  and risk ratio--in this case OR because using binomial link function.  

### logit/OR

```{r}

# dataset combining variables
d = cbind( hml3235, slept) %>%  as.data.frame()


  rstan_options(auto_write = TRUE)
  options(mc.cores = parallel::detectCores())
  
flist0 <- alist(
    hml3235 ~ dbinom( size=1 , prob=p ) ,
    logit(p) <- a + b * slept ,  # gives odds ratio
    c(a,b) ~ dnorm(0,10)  # regularization
)

fit0 = map( flist0 , data = d)

precis(fit0)@output

# odds
exp( precis(fit0)@output )  

# proportion-rel risk
# logisticTransform <- function(p) { exp(p)/(1 + exp(p)) }
# logisticTransform( precis(fit0)@output )  

```

### log-binomial/RR

```{r}

# dataset combining variables
d = cbind( hml3235, slept) %>%  as.data.frame()

library(rethinking, quietly = TRUE)

  rstan_options(auto_write = TRUE)
  options(mc.cores = parallel::detectCores())
  
flist0 <- alist(
    hml3235 ~ dbinom( size=1 , prob=p ) ,
    log(p) <- a + b * slept ,  # gives RR
    c(a,b) ~ dnorm(0, 10)  # regularization
)

fit0.map = map( flist0 , data = d, 
            start = list( a = log(.3) , b = log(.9) )  # .3, and .9 from glm model
            )

# precis(fit0)@output

# relative risk
exp( precis(fit0.map)@output )  

# convergence
# plot(fit0.map)

```


### more rethinking 

MCMC via STAN

```{r}
  
# when the quadratic approximate posterior is not a good fit, use MCMC via STAN
# takes about 25 minutes on macbook pro
fit0.stan <- map2stan( fit0 , data=d , warmup = 100, iter = 200 ) # default: warmup = 1e3, iter = 2e3

precis( fit0.stan )@output
exp( precis( fit0.stan )@output )

# basically--the same result


```



# Individual intercept for each survey (fit1)

primarily using lmer() for modeling effect 

### lmer

```{r}

fit1.lmer = glmer( hml3235 ~ slept + (1 | country_year_id ), family = binomial(link = "logit"),
                   data = d)

summary(fit1.lmer)
coef(fit1.lmer) 
exp(fixef(fit1.lmer))
```

### map

```{r}
Year = 1900 + floor( (i$hv008 - 1)/ 12 )  #convert CMC date to year
country_year = paste0( i$country, i$year )
country_year_id = factor( country_year )


d = data.frame( hml3235, slept, country_year_id) 

flist1 <- alist(
    hml3235 ~ dbinom( size=1 , prob=p ) ,
    logit(p) <- 
        a_country[country_year_id] +  
        b_slept * slept ,  
    a_country[country_year_id] ~ dnorm( 0 , 10 ),
    b_slept ~ dnorm(0, 10)  
)

fit1.map = map( flist1 , data = d, 
            start = list( 
                # a_country[country_year_id] = log(.3) , 
                b_slept = log(.9) )  # .3, and .9 from glm model
            )

saveRDS( fit1.map, file="fit1.map.rds")
readRDS( "fit1.map.rds" )

# relative risk
exp( precis(fit1.map)@output )  

# compare
compare( fit0.map, fit1.map )

# convergence
# plot(fit0.map)

```

## By country with # nets / household (fit1a)


```{r}
# Year = 1900 + floor( (i$hv008 - 1)/ 12 )  #convert CMC date to year
country_year = paste0( i$country, i$year )
country_year_id = factor( country_year )

nets_house = ih$nets_num_household

country_year = paste0( i$country, i$year )
country_year_id = factor( country_year )

nets_house = ih$nets_num_household
hist(nets_house)
hist(log(1+ nets_house))
hist( scale(nets_house))
hist( scale( log(1+ nets_house)) )

with(ih, table(slept, nets_num_household))
num_nets = ifelse( ih$nets_num_household >=13, 13, ih$nets_num_household )


d1a = data_frame( hml3235, slept, num_nets, country_year_id) 

```

### glm 

```{r}


#  fit 1 = variable intercept country-year (OR)
fit1a.glm = glm( hml3235 ~ country_year_id + slept + num_nets + slept*num_nets -1 , 
                data = d1a, 
                family = binomial(link = logit)
                )


### OR
exp(coef(fit1a.glm))

# exp( confint( fit1a.glm)) # slower , but more acurate

exp( confint.default( fit1a.glm))

# glimmer(hml3235 ~ paste0( x$iso3c, Year ) + slept -1, data = d, family = binomial )
```

### glmer

```{r}

fit1a.lmer = glmer( hml3235 ~ slept + num_nets +  slept * num_nets + (1 | country_year_id ), 
                    family = binomial(link = "logit"),
                   data = d1a)

# summary(fit1a.lmer)
# coef(fit1a.lmer) 
exp(fixef(fit1a.lmer))


```

### 'rethinking' post-analysis

```{r}


post = extract.samples( fit1a.lmer )
precis( exp(post), corr = TRUE)

slept.seq =  c(rep(0,14), rep(1,14))
num_nets.seq = rep(0:13, 2)

link.fit1a.lmer = function( slept, num_nets){ 
    post$Intercept + post$slept*slept + 
        post$num_nets * num_nets + 
        post$`slept:num_nets` * slept * num_nets
}

mu = mapply( link.fit1a.lmer, slept.seq, num_nets.seq)

pred.p = exp(apply( mu, 2, mean))
pred.PI = exp( apply( mu, 2, PI ))

pred = data.frame( slept.seq, num_nets.seq, pred.p, t( pred.PI ) )

ggplot( pred, aes(slept.seq, pred.p) ) + geom_line() + 
    geom_ribbon(aes( ymin = X5., ymax = X94.), alpha = .2) +
    scale_x_continuous( breaks = 0:1) +
    facet_wrap(~ num_nets.seq)
```

## By country with # person (fit1b)


```{r}
# Year = 1900 + floor( (i$hv008 - 1)/ 12 )  #convert CMC date to year
country_year = paste0( i$country, i$year )
country_year_id = factor( country_year )

household_num = ih$household_num

hist(household_num)
hist(log(1+ household_num))
hist( scale(household_num))
hist( scale( log(1+ household_num)) )

with(ih, table(slept, household_num))

# create number of persons categories
num_persons.f = cut( 
    household_num, breaks = c( seq(1, 34, 3), Inf), 
    include.lowest = TRUE , right = FALSE
    ) # summary(num_persons.f)

num_persons = as.numeric( num_persons.f )  # convert to numeric

# use factor levels to interpret ranges, e.g. levels( num_persons.f)[4]

d1b = data_frame( hml3235, slept, num_persons, country_year_id) 

```

### glm

```{r}

# Year = 1900 + floor( (i$hv008 - 1)/ 12 )  #convert CMC date to year
country_year = paste0( i$country, i$year )
country_year_id = factor( country_year )

nets_person = ih$nets_num_household / ih$household_num

d = data_frame( hml3235, slept, num_persons.f, country_year_id) 

#  fit 1 = variable intercept country-year (OR)
fit1b.glm = glm( hml3235 ~ country_year_id + slept + num_persons.f + slept*num_persons.f -1 , 
                data = d, 
                family = binomial(link = logit)
                )


### OR
exp(coef(fit1b.glm))

exp( confint.default( fit1b.glm))

# glimmer(hml3235 ~ paste0( x$iso3c, Year ) + slept -1, data = d, family = binomial )
```

### glmer

```{r}

fit1b.lmer = glmer( hml3235 ~ slept + 
                        num_persons + 
                        slept * num_persons + 
                        (1 | country_year_id ), 
                    family = binomial(link = "logit"),
                   data = d1b)

# summary(fit1b.lmer)
# coef(fit1b.lmer) 
exp(fixef(fit1b.lmer))

```

### 'rethinking' post-analysis

```{r}

post = extract.samples( fit1b.lmer )
precis( exp(post), corr = TRUE)

slept.seq =  c(rep(0,12), rep(1,12))
num_persons.seq = rep( 1:12, 2)

post = extract.samples( fit1b.lmer )

link.fit1b.lmer = function( slept, num_persons){ 
    post$Intercept +
        post$slept * slept + 
        post$num_persons * num_persons + 
        post$`slept:num_persons` * slept * num_persons
}
mu = mapply( link.fit1b.lmer, slept.seq, num_persons.seq)

o2p = function(x) x/(1+x) # convert odds to probaility (after converting log-odds to odds)
pred.p = o2p( exp( apply( mu, 2, mean) ) )
pred.PI = o2p( exp( apply( mu, 2, PI ) ) )

pred = data.frame( slept.seq, num_persons.seq, pred.p, t( pred.PI ) )

# calc RR
num_persons.seq = 1:12

linkRR.fit1b.lmer = function( num_persons){ 
    sleep0 = post$Intercept +
        post$slept * 0 + 
        post$num_persons * num_persons + 
        post$`slept:num_persons` * 0 * num_persons
    
    sleep1 = post$Intercept +
        post$slept * 1 + 
        post$num_persons * num_persons + 
        post$`slept:num_persons` * 1 * num_persons
    
    RR = o2p( exp( sleep1 ) )  /  o2p( exp( sleep0 ) )
    
    return(RR)
}
mu.rr = mapply( linkRR.fit1b.lmer, num_persons.seq)
pred.rr = apply( mu.rr, 2, mean) 
pred.rr.PI = apply( mu.rr, 2, PI )
rr = data.frame( num_persons.seq, pred.rr, t( pred.rr.PI ))
rr$label = paste( round(rr$pred.rr,2) ,"[", round(rr$X5.,2) ,"-", round(rr$X94.,2), "]" )

num_persons_labeller = function(variable,value) levels(num_persons.f)[value] 

ggplot( pred, aes(slept.seq, pred.p)) + 
    geom_line() + 
    geom_ribbon(aes( ymin = X5., ymax = X94.), alpha = .2) +
    geom_label( data = rr, x=.5, y=.4, aes(label = label)) +
    scale_x_continuous( breaks = 0:1) +
    facet_wrap(~ num_persons.seq, labeller = num_persons_labeller ) +
    ggtitle("Effect of sleeping under net", subtitle = "by number of persons in household")

```


## By country with # nets and # persons (fit1c)

### glmer

```{r}

d = data_frame( hml3235, slept, num_nets, num_persons, country_year_id) 


fit1c.lmer = glmer( hml3235 ~ slept + 
                        num_nets + 
                        num_persons +  
                        slept * num_nets +
                        slept * num_persons +
                        (1 | country_year_id ), 
                    family = binomial(link = "logit"),
                   data = d)

# summary(fit1c.lmer)
# coef(fit1c.lmer) 
exp(fixef(fit1c.lmer))

```

### 'rethinking' post-analysis

```{r}

post = extract.samples( fit1c.lmer )
precis( exp(post), corr = TRUE)

grid = expand.grid( slept.seq = 0:1, 
                    num_nets.seq = 0:7, 
                    num_persons.seq = 1:12)

slept.seq =  grid[[1]]
num_nets.seq = grid[[2]]
num_persons.seq = grid[[3]]

link.fit1c.lmer = function( slept, num_nets, num_persons){ 
    post$Intercept +
        post$slept * slept + 
        post$num_nets * num_nets + 
        post$num_persons * num_persons +
        post$`slept:num_nets` * slept * num_nets +
        post$`slept:num_persons`  * slept * num_persons
}
mu = mapply( link.fit1c.lmer, slept.seq, num_nets.seq, num_persons.seq)

pred.p = exp(apply( mu, 2, mean))
pred.PI = exp(apply( mu, 2, PI ))

pred = data.frame( slept.seq, num_nets.seq, num_persons.seq, pred.p, t( pred.PI ) )

ggplot( pred, aes(slept.seq, pred.p)) + geom_line() + 
    geom_ribbon(aes( ymin = X5., ymax = X94.), alpha = .2) +
    scale_x_continuous( breaks = 0:1) +
    facet_grid( num_persons.seq ~ num_nets.seq )

```

### compare models

```{r}
compare( fit1a.lmer , fit1b.lmer )

```


## By country with wealth index (fit1d)

### glmer

```{r}

count(ih, hv270)

wealth = ih$hv270

d = data_frame( hml3235, slept, wealth, country_year_id) 


fit1d.lmer = glmer( hml3235 ~ slept + 
                        wealth + 
                        slept * wealth +
                        (1 | country_year_id ), 
                    family = binomial(link = "logit"),
                   data = d)

# summary(fit1c.lmer)
# coef(fit1c.lmer) 
exp(fixef(fit1d.lmer))

```

### 'rethinking' post-analysis

```{r}

post = extract.samples( fit1d.lmer )

precis( exp(post), corr = TRUE)

grid = expand.grid( slept.seq = 0:1, 
                    wealth.seq = 1:5)

slept.seq =  grid[[1]]
wealth.seq = grid[[2]]

link.fit1d.lmer = function( slept, wealth){ 
    post$Intercept +
        post$slept * slept + 
        post$wealth * wealth + 
        post$`slept:wealth`  * slept * wealth
}
mu = mapply( link.fit1d.lmer, slept.seq, wealth.seq)

pred.p = exp(apply( mu, 2, mean))
pred.PI = exp(apply( mu, 2, PI ))

pred = data.frame( slept.seq, wealth.seq, pred.p, t( pred.PI ) )

# calc RR
num_persons.seq = 1:12

linkRR.fit1d.lmer = function( wealth){ 
    sleep0 = post$Intercept +
        post$slept * 0 + 
        post$wealth * wealth + 
        post$`slept:wealth` * 0 * wealth
    
    sleep1 = post$Intercept +
        post$slept * 1 + 
        post$wealth * wealth + 
        post$`slept:wealth` * 1 * wealth
    
    RR = o2p( exp( sleep1 ) )  /  o2p( exp( sleep0 ) )
    
    return(RR)
}
mu.rr = mapply( linkRR.fit1d.lmer, wealth.seq)
pred.rr = apply( mu.rr, 2, mean) 
pred.rr.PI = apply( mu.rr, 2, PI )
rr = data.frame( wealth.seq, pred.rr, t( pred.rr.PI ))
rr$label = paste( round(rr$pred.rr,2) ,"[", round(rr$X5.,2) ,"-", round(rr$X94.,2), "]" )

ggplot( pred, aes(slept.seq, pred.p)) + 
    geom_line() + 
    geom_ribbon(aes( ymin = X5., ymax = X94.), alpha = .2) +
    geom_label( data = rr, x=.5, y=.4, aes(label = label)) +
    scale_x_continuous( breaks = 0:1) +
    facet_grid(. ~ wealth.seq) +
    ggtitle("Effect of sleeping under net", subtitle = "by wealth quintile (varying intercept)")

```




# Varying slopes (fit2)
## By country with wealth index (fit2d)

### glmer

```{r}

count(ih, hv270)

wealth = ih$hv270

d = data_frame( hml3235, slept, wealth, country_year_id) 


fit2d.lmer = glmer( hml3235 ~ slept + 
                        wealth + 
                        slept * wealth +
                        (1 + slept + wealth| country_year_id ), 
                    family = binomial(link = "logit"),
                   data = d)

# summary(fit2d.lmer)
# coef(fit2d.lmer) 
exp(fixef(fit2d.lmer))

# compare with model 1d
# (Intercept)        slept       wealth slept:wealth 
#    1.0446968    0.9842546    0.6659320    0.9761877 

```

### 'rethinking' post-analysis

```{r}

post = extract.samples( fit2d.lmer )

precis( exp(post), corr = TRUE)

grid = expand.grid( slept.seq = 0:1, 
                    wealth.seq = 1:5)

slept.seq =  grid[[1]]
wealth.seq = grid[[2]]

link.fit2d.lmer = function( slept, wealth){ 
    post$Intercept +
        post$slept * slept + 
        post$wealth * wealth + 
        post$`slept:wealth`  * slept * wealth
}
mu = mapply( link.fit1d.lmer, slept.seq, wealth.seq)

pred.p = exp(apply( mu, 2, mean))
pred.PI = exp(apply( mu, 2, PI ))

pred = data.frame( slept.seq, wealth.seq, pred.p, t( pred.PI ) )

# calc RR
num_persons.seq = 1:12

linkRR.fit2d.lmer = function( wealth){ 
    sleep0 = post$Intercept +
        post$slept * 0 + 
        post$wealth * wealth + 
        post$`slept:wealth` * 0 * wealth
    
    sleep1 = post$Intercept +
        post$slept * 1 + 
        post$wealth * wealth + 
        post$`slept:wealth` * 1 * wealth
    
    RR = o2p( exp( sleep1 ) )  /  o2p( exp( sleep0 ) )
    
    return(RR)
}
mu.rr = mapply( linkRR.fit2d.lmer, wealth.seq)
pred.rr = apply( mu.rr, 2, mean) 
pred.rr.PI = apply( mu.rr, 2, PI )
rr = data.frame( wealth.seq, pred.rr, t( pred.rr.PI ))
rr$label = paste0( round(rr$pred.rr,2) ,"\n[", round(rr$X5.,2) ,"-", round(rr$X94.,2), "]" )

ggplot( pred, aes(slept.seq, pred.p)) + 
    geom_line() + 
    geom_ribbon(aes( ymin = X5., ymax = X94.), alpha = .2) +
    geom_label( data = rr, x=.5, y=.9, aes(label = label)) +
    scale_x_continuous( breaks = 0:1) +
    facet_grid(. ~ wealth.seq) +
    ggtitle("Effect of sleeping under net", subtitle = "by wealth quintile (varying intercept and slope.  fit2d.)")

```



# compare models

```{r}
compare( fit1a.lmer , fit1b.lmer )

```


## GLM by cluster (fit2)

```{r}

Year = 1900 + floor( (i$hv008 - 1)/ 12 )  #convert CMC date to year
country_year_cluster = paste0( i$country, i$year, i$hv001 )
country_year_cluster_id = factor( country_year_cluster )

d = data.frame( cbind( hml3235, slept, country_year_cluster) )


#  fit 2 = variable intercept country-year-cluster (OR)
fit2.glm = glm( hml3235 ~ country_year_cluster_id + slept -1 , 
                      data = d, 
                      family = binomial(link = logit))

### OR
exp(coef(fit2.glm))


# glimmer(hml3235 ~ paste0( x$iso3c, Year ) + slept -1, data = d, family = binomial )
```

### Map2Stan version of individual intercepts for each survey

```{r}
# logit
m.country <- alist(
    hml3235 ~ dbinom( prob=p , size=1 ) ,
    
    logit(p) <- 
        a +
        a_survey[country_year_id] +  
        b_slept * slept ,  
    
    a ~ dnorm(0, 10),
    a_survey[country_year_id] ~ dnorm( 0 , 10 ),
    b_slept ~ dnorm(0, 10) 
)

fit.m.country = map( m.country , data = d, 
                     start = list( a = log(.3) , a_survey = log(1),
                                   b_slept = log(1) ) )

save(d,fit.m.country, file = "fit.m.country.rda")
load("fit.m.country.rda")

exp( precis( fit.m.country, depth = 2 )@output )

#@ log
m.country.log <- alist(
    hml3235 ~ dbinom(  size=1 , prob=p ) ,
    
    log(p) <- 
        a +
        a_survey[country_year_id] +  
        b_slept * slept ,  
    
    a ~ dnorm(0, 1),
    a_survey[country_year_id] ~ dnorm( 0 , 10 ),
    b_slept ~ dnorm(0, 1) 
)

fit.m.country.log = map2stan( m.country.log , data = d, iter = 200)
# fit.m.country.log = map( m.country.log , data = d  , 
#                      start = list( a = log(.3) , a_survey = log(1),
#                                    b_slept = log(1) ) )
plot(fit.m.country.log)

save(d,fit.m.country.log, file = "fit.m.country.log.rda")
load("fit.m.country.log.rda")

exp( precis( fit.m.country.log, depth = 2 )@output )

# refit with fewer iterations...
fit.m.country.log.short = map2stan( m.country.log , data = d, iter = 200)

# model fitting and convergence:
# compare(fit.m.country.log, ...)
post = extract.samples(fit.m.country.log )
pred = predict(fit.m.country.log )

x = rep( 0:1, 1000)
y = post$a + x*post$b_slept
plot(0,0, type = "n", xlab="slept under net", ylab = "parasitemia", ylim = 0:1,
     xaxt = "n", xlim = 1:2
)
axis( 1, at = 1:2, labels = c("did not", "did") )

```







## Multi-level, median a postiori rethinking package

estimates both risk (intercept)  and risk ratio--in this case OR because using binomial link function.  

```{r}


  #calc year-month from CMC date
Year = 1900 + floor( (x$hv008 - 1)/ 12 )
Month = x$hv008 - 12 * (Year - 1900)

country = coerce_index( x$iso3c )

urban_rural = x$hv025

psu = paste0( x$iso3c, Year, x$hv001  ) 
psu_id = coerce_index( psu )

household = paste0( x$iso3c, Year, x$hv001, x$hv002 ) 
household_id = coerce_index( household )

# ML0 fit country
d = as.data.frame( cbind( hml3235, slept, country) )
ml0 <- alist(
    hml3235 ~ dbinom( prob=p , size=1 ) ,
    logit(p) <- 
        a_country[country] +  
        b_slept * slept ,  
    a_country[country] ~ dnorm( 0 , 10 ),
    b_slept ~ dnorm(0, 10) 
)
fitml0 = map( ml0 , data = d)
exp( precis(fitml0)@output )

# fit country and urban/rural
d = as.data.frame( cbind( hml3235, slept, country, urban_rural ) )

ml0.1 <- alist(
    hml3235 ~ dbinom( prob=p , size=1 ) ,
    logit(p) <- a + 
        a_country[country] +  
        a_urban[urban_rural] +
        b_slept * slept ,  
    a_country[country] ~ dnorm( 0 , 10 ),
    a_urban[urban_rural] ~ dnorm( 0 , 10 ),
    c(a, b_slept) ~ dnorm(0, 10) 
)
fitml0.1 = map( ml0.1 , data = d)
exp( precis(fitml0.1)@output )

compare(fitml0, fitml0.1)

# Map2Stan  ####
d = as.data.frame( cbind( hml3235, slept, country) )
ml1 <- alist(
    hml3235 ~ dbinom(1, p) ,
    logit(p) <- a + 
        a_country[country] +  
        # a_psu[psu_id] +  
        # a_household[household_id] + 
        b_slept * slept ,  
    
    a_country[country] ~ dnorm( 0 , sigma_country ),
    # a_psu[psu_id] ~ dnorm( 0 , sigma_psu ),
    # a_household[household_id] ~ dnorm( 0 , sigma_household ),
    
    c(a, b_slept) ~ dnorm(0, 10) ,
    
    sigma_country ~ dcauchy(0,1)
    # sigma_psu ~ dcauchy(0,1)
    # sigma_household ~ dcauchy(0,1)
)

# first fit 
fit_ml1 = map2stan( ml1 , data = d, chains = 1, cores = 1) 
exp( precis(fit_ml1, depth = 1 )@output )

# full fit 
fit_ml1.full  = map2stan( fit_ml1 , data = d , chains = 4, cores = 2) 
exp( precis(fit_ml1.full, depth = 1 )@output)

saveRDS(fit_ml1.full , "fit_ml1.full.rds")

# m2 <- map2stan(  # TAKES 10 minutes to run!
#     alist(
#         hml3235 ~ dbinom(1, theta),
#         logit(theta) = a + bs * slept ,
#         a ~ dnorm(0,10),
#         bs ~ dnorm(0,10)
#     ) ,
#     data=d, chains=2, cores=1 )
# 
# exp( precis(m2)@output )

# library(arm)
# lmer( hml3235 ~ slept , data = d)

```


## Visualize parasitemia versus bednet coverage by cluster

Because the samples are paired, need to be sure that each part of the pair exists.  For a given level on x-axis, there may be only groups from high prevalence areas that all use nets, so there is no no-net group, that get compared with groups from low-prevalence areas where there are none in the use-net group.  So, the loess line is comparing groups using nets in high prevalence areas to groups not using nets in low prevalence areas; both groups may have roughly same level of risk, therefore it looks as if the nets are not protective.

Possible solutions:  
1. make sure that each cluster has a parasitemia rate for no-net and net group (filter on complete.cases() )


## Group by cluster (hv001)

```{r}
 d1  = x %>% 
  mutate( 
    
    parasitemia = ifelse( hml32 %in% 1 | hml35 %in% 1, 1, ifelse( hml32 %in% 0 | hml35 %in% 0, 0, NA) ),
    
    know_parasitemia = parasitemia %in% 0:1 ,
    
    know_net_use =  hml19 %in% 0:1 | hml20 %in% 0:1 ,
    
    used_net = hml19 %in% 1 | hml20 %in% 1
    
          ) %>%
  
  group_by( country, survey, year, hv001) %>%
  
  summarise(
    
    pfpr = sum( parasitemia, na.rm = TRUE ) / sum( know_parasitemia, na.rm = TRUE ),
    
    slept_under_net = sum( used_net, na.rm = TRUE ) / sum( know_net_use, na.rm = TRUE ),
    
    # net_per_person = sum( hml1, na.rm  = TRUE ) / sum( hv012, na.rm  = TRUE ),
    
    pfpr_net = sum( parasitemia & used_net, na.rm = TRUE ) / sum( know_parasitemia & used_net, na.rm = TRUE ),
    
    pfpr_no_net = sum( parasitemia & !used_net, na.rm = TRUE ) / sum( know_parasitemia & !used_net, na.rm = TRUE ),
    
    n = n(),
    n_para = sum( know_parasitemia, na.rm = TRUE ),
    n_net_para = sum( know_parasitemia & used_net, na.rm = TRUE ) ,
    n_no_net = sum( know_parasitemia & !used_net, na.rm = TRUE )
  ) 

```

```{r}
ggplot( d1, aes( x = slept_under_net, y = pfpr)) +
  geom_point() +
  geom_smooth()

```

