---
title: "DHS Datasets"
author: "jp"
date: "10/5/2016"
output:
  html_notebook: default
  pdf_document: default
---

```{r, include= FALSE}

library(knitr)
library(scales, quietly = TRUE)
library(data.table)
library(dtplyr)
library(tidyverse, quietly = TRUE)
library(countrycode)

opts_chunk$set( echo = TRUE, cache = TRUE )

# location of DHS folder.  include DHS files as: paste0( dhs, <dhs file here>)
dhs = "../DHS/"
```

# Survey datasets

## Select country/surveys to include 

```{r}

source( paste0( dhs, "file_list.r") )

subsahara = countrycode_data %>%
  filter( region %in% c("Middle Africa", "Western Africa", "Eastern Africa", "Southern Africa")) %>% 
  select( country.name, iso3c ) %>%
  filter( !country.name %in% "Lesotho") 


files = files %>% 
  mutate( country.name = countrycode(country, "country.name", "country.name")) %>%
  filter( survey!="", year >= 2000 , country.name %in% subsahara$country.name)

# View(files)

# not_needed = c('Bangladesh', 'Pakistan', 'Philippines', 'Comoros')

survey_list = files %>% count(country, survey, year) %>% select(-n)
# View(survey_list)

nsurveys = nrow(survey_list)
ncountries = length( unique( files$country ) )

cat( "There are", nsurveys, "surveys since 2000 from",  ncountries ,"countries in Subsaharan Africa.")

# kable( survey_list )
```

## Surveys with parasitemia biomarkers

```{r}
dd <-readRDS( paste0( dhs, "dictionaryNew.rds")  )  # about 6 sec

parasitemia_variables = c("hml32", "hml35")

parasitemia_surveys = dd %>% 
  filter( survey != "", year >= 2000 , iso3c %in% subsahara$iso3c ) %>%
  filter( Item_Name %in% toupper( parasitemia_variables) ,
          value %in% 0:1 ) %>% # comment this line out to get full list of variables
  filter( !grepl( "NA", Item_Label, ignore.case = FALSE, fixed = TRUE)  ) %>%
        select( Item_Name, country, survey, value, year) %>%
      count( country, survey, year, Item_Name) %>%
    mutate( 
        has = ifelse( n > 0 , "Y", "N" ) 
        ) %>%
    select( -n) %>%
  spread( Item_Name ,  has, fill = "N" )

 kable(parasitemia_surveys, row.names = TRUE)  
  
```

## Household member dataset

## Select **individual** variables of interest for bednet study and check to see which survey file has the values

```{r}

individual_vars  = c( 
        'hml32', 'hml35', # RDT and micrscopy
        'hml1', "hml4", "hml5", "hml6", "hml7", "hml8",  # bednets
        'hml19', 'hml20', 'hml12', # slept under llin
        'hml16', # corrected age in years
        "hv253", # dwelling sprayed (some multiple responses)
        "hv001", "hv002", 'HV003', 'HHID', 'HVIDX',
        "HV008" # date of interview
       )
  
variables <- unique( toupper( individual_vars) )
variables = variables[ order(variables) ]

# translation table of dhs file codes to file names
filecodes = data_frame( abrev  = c("kr", "br", "ir", "pr", "hr", "hr" ), 
                            file = c("Children's Recode","Supplemental Births Recode",
                                     "Individual Recode",
                                     "Household Member Recode", "Household Recode", 
                                     "Supplemental Household Recode")) 

 
# list variables and the files where each is found
survey_files = dd %>% semi_join( parasitemia_surveys ) %>%
    filter( Item_Name %in% variables ) %>%
    filter( !grepl( "NA", Item_Label, ignore.case = FALSE, fixed = TRUE) ,
            !grepl( "Presence", Item_Label, ignore.case = FALSE, fixed = TRUE)
            ) %>%
    left_join( filecodes , by = "file", copy = TRUE) %>%
    mutate( 
        abrev = ifelse( is.na(abrev) , tolower( substr(file, 3, 4) ) , abrev)
        ) %>%
    left_join( filecodes , by = "abrev", copy = TRUE) %>%
    rename( file = file.y ) %>%
    count( Item_Name, Item_Label, country, survey, year, file ) %>%
    mutate( has = ifelse( n > 0, "Y", NA )) %>% 
    ungroup() %>%
    count( Item_Name, file ) %>%
    spread( file, nn, fill = "N" )

# View(dataset_surveys)

kable(survey_files)
```

## variable definitions...

Select the first available definition from dictionary.  Note that some variables may be defined differently for some countries, which will not be apparent here. 

```{r}
variable_defn = dd %>% semi_join( parasitemia_surveys, by = c("country", "survey", "year") ) %>%
    filter( Item_Name %in% variables ) %>%
    filter( !grepl( "NA", Item_Label, ignore.case = FALSE, fixed = TRUE) ,
            !grepl( "Presence", Item_Label, ignore.case = FALSE, fixed = TRUE)
            ) %>%
    left_join( filecodes , by = "file", copy = TRUE) %>%
    mutate( 
        abrev = ifelse( is.na(abrev) , tolower( substr(file, 3, 4) ) , abrev)
        ) %>%
    left_join( filecodes , by = "abrev", copy = TRUE) %>%
    rename( file = file.y ) %>%
    filter( grepl( "member", file, ignore.case= T)) %>%
    count( Item_Name, Item_Label) %>%
    select( -n) %>%
    arrange( Item_Label ) %>%
    as.data.frame()

variable_defn
```


## create household member dataset. run one time after updating variables or adding surveys

```{r, eval=FALSE, include = FALSE}
source( paste0( dhs, "getSurveyGIS.R") )
source( "load_survey_file.R") 

nsurveys = nrow( parasitemia_surveys )

parasitemia_surveys = arrange( parasitemia_surveys, desc(year) )
index = row.names(parasitemia_surveys)

p <- progress_estimated(nsurveys)

x <- list()

translate = FALSE  # optional translation of values to label.  
  # NB: Does not work well for all variables

for (i in seq_along( index )  ){
  
   p$pause(0.1)$tick()$print()
  
  .country = parasitemia_surveys[i, ]$country
  .iso3c = countrycode( parasitemia_surveys[i, ]$country, "country.name", "iso3c")
  .survey = parasitemia_surveys[i, ]$survey
  .year = parasitemia_surveys[i, ]$year
  .file = "Household Member Recode"
  .survey_year = paste(.survey, .year)

  print( paste(.country, .survey, .year, .file) )
  
  x[[i]] <- try( silent = TRUE,
    load_survey_file( design = FALSE,
                        # printout = TRUE ,  # un comment when testing
                        vars = variables, 
                        .country = .country, 
                        .survey = .survey,
                        .year = .year,
                        .file = .file  )
  )
  
      
    if ( class(x[[i]]) == "try-error" )  next()
     

  x[[i]]$country = rep( .country, nrow(x[[i]]) )
  x[[i]]$iso3c = rep( .iso3c, nrow(x[[i]]) )
  x[[i]]$year = rep( .year, nrow(x[[i]]) )
  x[[i]]$survey = rep( .survey, nrow(x[[i]]) )

  # TODO: optional- relate dictionary to survey                    )

  # # TODO: only for columns with >4 values (others easier to keep as numeric)
  # # Deriving this table should happen once, before loop
  # v = dd %>% filter( tolower(Item_Name) %in% variables) %>% group_by( Item_Name, label ) %>% 
  #   summarise( n = n()) %>% group_by( Item_Name) %>% count(Item_Name) %>% filter(nn>4) 

  if (translate){
    
      # refine dictionary for this country-survey 
      dict = dd %>% filter_( ~country == .country , ~survey == .survey, ~year == .year,
                        ~Item_Name %in% toupper(names(has_vars))
      )
     
    y = x[[i]]
  
    for (col in 1:length(colnames(y))){
      column = tolower( colnames(y)[col] )
      
      dict_var = dict %>% 
        mutate( item_name = tolower(Item_Name)) %>%
        filter_( ~item_name == column ) %>% # , ~file == "Household Member Recode"
        count( value, label) %>% select(-n)
      
      old_value = y[, col]
      new_value = dict_var[ match( old_value, dict_var$value ), ]$label
      y[, col] = ifelse( is.na( new_value), old_value, new_value)
    }
      x[[i]] = y
}

  # if (is.null(X) ){ X <- x; next()}
  # X = rbindlist( list(X, x), use.names = TRUE, fill = TRUE)
  
}
  
# X = dplyr:: bind_rows( x )  # will throw error is column types dont match 
X = rbindlist(x, fill = TRUE) 

#TODO FILTER RECORDS--apply filter, like keeping only children with parasitemia, is not unnecessarily large
  
 # if (!is.na( subset_filter ))  X = X %>% filter_( subset_filter )

 saveRDS(X, "household_member_bednet.rds")
 # object.size(X)

```

## create individual file by filtering household member file to those children with parasitemia result

```{r, eval= FALSE }
hm = readRDS( "household_member_bednet.rds"  ) %>% as_tibble()

 # These countries were supposed to have results, but there are none...
hm %>% 
    mutate(
        RDT = hml32 %in% 0:1 ,
        slide = hml35 %in% 0:1 
    ) %>%
    count( country, survey, year, RDT, slide ) %>% ungroup %>%
    count( country, survey, year ) %>% filter(nn == 1)
    
    
# filter to those with parsitemia result

has.test.result = hm$hml32 %in% 0:1 | hm$hml35 %in% 0:1
i = hm[ has.test.result ,  ]

cat( "There are", nrow(i), "children with a either a positve or negative parsitemia result from",
     count(i, country, survey, year) %>% nrow, "surveys in ",
     count(i, country) %>% nrow, "countries" 
     )

saveRDS(i, "individual_bednet.rds" )

```

## Create household dataset by aggregating household member file 

net variables:
- hml1: number of nets
- hml4: months ago obtained net
- hml5: was net treated with insecticide when obtained (binary)
- hml6: Mosquito net treated with insecticide (6 options)
- hml7: brand of bednet
- hml8: bednet treated since obtaining

```{r, eval=FALSE}

hm = readRDS( "household_member_bednet.rds"  ) %>% as_tibble()

members = hm %>% 
    count( country, iso3c, survey, year, hv001, hv002) %>%
    rename( household_num = n ) %>%
    ungroup

# calculate avg household
# h %>% group_by(country, iso3c, survey, year) %>% summarise( 
#     avg_household = mean(n) 
#     )

nets = hm %>% 
    group_by( country, iso3c, survey, year, hv001, hv002) %>%
    summarise( nets_num = max(hml1, na.rm = TRUE) ) %>%
    ungroup

h = inner_join( members, nets, 
                by = c("country", "iso3c", "survey", "year", "hv001", "hv002"))

cat("there are", nrow(h), "households")

saveRDS( h, file = "household_bednet.rds" )
```


# Load datasets and see which surveys are included


```{r }

i = readRDS(  "individual_bednet.rds" ) %>% as_tibble()
h = readRDS(  "household_bednet.rds" ) %>% as_tibble()

# v001 & hv001 - cluster number
# v002 & hv002 - household number
# b16 & hvidx - line number of child in household (for linking moms and kids)

#join childrens to household to get # persons per household, then join that with household member file.

ih = inner_join( i, h, by = c( "country" = "country",
                               "iso3c"="iso3c",
                              "year" = "year",
                              "survey" = "survey",
                              "hv001" = "hv001",
                              "hv002" = "hv002")
)

data_ncountries = length( unique(ih$country))
data_nsurveys = nrow( count( ih, country, survey, year)  )
nvars = ncol(ih)
rows = nrow(ih)

cat( "The combined individual/household file includes", data_nsurveys, "surveys from", data_ncountries, "countries with", nvars, "columns, ", rows, "rows and is " , comma(object.size(ih)[1]), "bytes")

# list surveys included
kable( ih %>% count( country, survey, year ) , row.names = TRUE)
```

# Individual risk-ratio for bednets (fit0)

```{r}



## Slept under net
# filter( dd, Item_Name == "HML12") %>% count(value, label)
slept = integer( nrow(i))
slept[ i$hml12 %in% 1:5] = 1

ts = table(slept, useNA = 'always') 
ts

p = function(n0, yes){ yes/(n0 + yes)}

p(ts[1], ts[2])



## is parasitemic by rdt or slide
hml3235 = integer( nrow(i))
hml3235[ i$hml32 == 1 | i$hml35 == 1 ] = 1

tp = table(hml3235, useNA = 'always')
tp
p(tp[1], tp[2])


# dataset combining variables
d = cbind( hml3235, slept) %>%  as.data.frame()

# tally
count(d, hml3235, slept) 



```


## epitools package

estimates crude RR 

```{r}
library(epitools)
r = riskratio(d$slept, d$hml3235)  
# str(r)
r$measure[2,]

```

## GLM logistic (OR) 

GLM gives estimate of overall risk, and risk ratio

```{r}
library(base)
# logistic regression for rr
glm.logistic = glm( hml3235 ~ slept , data = d, family = binomial(link = "logit"))

### Risk (intercept) and odds ratios 
exp(coef(glm.logistic))

```


### classical log-binomial model (RR)

```{r}
# dataset combining variables
d = cbind( hml3235, slept) %>%  as.data.frame()


# log-binomial regression for rr
glm.log.binomial = glm( hml3235 ~ slept  , data = d, family = binomial(link = "log"))

### Risk (intercept) and risk ratios (under poisson assumption)
exp(coef(glm.log.binomial))

```


## Bayesian, maximum a postiori (MAP) rethinking package

estimates both risk (intercept)  and risk ratio--in this case OR because using binomial link function.  

### logit/OR

```{r}

# dataset combining variables
d = cbind( hml3235, slept) %>%  as.data.frame()

library(rethinking, quietly = TRUE)

  rstan_options(auto_write = TRUE)
  options(mc.cores = parallel::detectCores())
  
flist0 <- alist(
    hml3235 ~ dbinom( size=1 , prob=p ) ,
    logit(p) <- a + b * slept ,  # gives odds ratio
    c(a,b) ~ dnorm(0,10)  # regularization
)

fit0 = map( flist0 , data = d)

precis(fit0)@output

# odds
exp( precis(fit0)@output )  

# proportion-rel risk
# logisticTransform <- function(p) { exp(p)/(1 + exp(p)) }
# logisticTransform( precis(fit0)@output )  

```

### log-binomial/RR

```{r}

# dataset combining variables
d = cbind( hml3235, slept) %>%  as.data.frame()

library(rethinking, quietly = TRUE)

  rstan_options(auto_write = TRUE)
  options(mc.cores = parallel::detectCores())
  
flist0 <- alist(
    hml3235 ~ dbinom( size=1 , prob=p ) ,
    log(p) <- a + b * slept ,  # gives odds ratio
    c(a,b) ~ dnorm(0, 10)  # regularization
)

fit0 = map( flist0 , data = d, 
            start = list( a = log(.01) , b = log(1) )  # .3, and .9 from glm model
            )

# precis(fit0)@output

# relative risk
exp( precis(fit0)@output )  

# convergence
plot(fit0)

```


### more rethinking 

MCMC via STAN

```{r}
  
# when the quadratic approximate posterior is not a good fit, use MCMC via STAN
# takes about 25 minutes on macbook pro
fit0.stan <- map2stan( fit0 , data=d , warmup = 100, iter = 200 ) # default: warmup = 1e3, iter = 2e3

precis( fit0.stan )@output
exp( precis( fit0.stan )@output )

# basically--the same result


```


# Individual intercept for each survey
## GLM by country (fit1)
```{r}

Year = 1900 + floor( (i$hv008 - 1)/ 12 )  #convert CMC date to year
country_year = paste0( i$country, i$year )
country_year_id = factor( country_year )


d = data_frame( hml3235, slept, country_year_id) 


#  fit 1 = variable intercept country-year (OR)
fit1.glm = glm( hml3235 ~ country_year_id + slept -1 , 
                data = d, 
                family = binomial(link = logit)
                )


### OR
exp(coef(fit1.glm))


# glimmer(hml3235 ~ paste0( x$iso3c, Year ) + slept -1, data = d, family = binomial )
```


## GLM by country with # nets / household (fit1a)


```{r}

# Year = 1900 + floor( (i$hv008 - 1)/ 12 )  #convert CMC date to year
country_year = paste0( i$country, i$year )
country_year_id = factor( country_year )

nets_house = ih$nets_num

d = data_frame( hml3235, slept, nets_house, country_year_id) 


#  fit 1 = variable intercept country-year (OR)
fit1a.glm = glm( hml3235 ~ country_year_id + slept + nets_house + slept*nets_house -1 , 
                data = d, 
                family = binomial(link = logit)
                )


### OR
exp(coef(fit1a.glm))

# exp( confint( fit1a.glm)) # slower , but more acurate

exp( confint.default( fit1a.glm))

# glimmer(hml3235 ~ paste0( x$iso3c, Year ) + slept -1, data = d, family = binomial )
```

## GLM by country with # nets / person (fit1b)


```{r}

# Year = 1900 + floor( (i$hv008 - 1)/ 12 )  #convert CMC date to year
country_year = paste0( i$country, i$year )
country_year_id = factor( country_year )

nets_person = ih$nets_num / ih$household_num

d = data_frame( hml3235, slept, nets_house, country_year_id) 


#  fit 1 = variable intercept country-year (OR)
fit1b.glm = glm( hml3235 ~ country_year_id + slept + nets_person + slept*nets_person -1 , 
                data = d, 
                family = binomial(link = logit)
                )


### OR
exp(coef(fit1b.glm))

exp( confint.default( fit1b.glm))

# glimmer(hml3235 ~ paste0( x$iso3c, Year ) + slept -1, data = d, family = binomial )
```

## GLM by cluster (fit2)

```{r}

Year = 1900 + floor( (i$hv008 - 1)/ 12 )  #convert CMC date to year
country_year_cluster = paste0( i$country, i$year, i$hv001 )
country_year_cluster_id = factor( country_year_cluster )

d = data.frame( cbind( hml3235, slept, country_year_cluster) )


#  fit 2 = variable intercept country-year-cluster (OR)
fit2.glm = glm( hml3235 ~ country_year_cluster_id + slept -1 , 
                      data = d, 
                      family = binomial(link = logit))

### OR
exp(coef(fit2.glm))


# glimmer(hml3235 ~ paste0( x$iso3c, Year ) + slept -1, data = d, family = binomial )
```

### Map2Stan version of individual intercepts for each survey

```{r}
# logit
m.country <- alist(
    hml3235 ~ dbinom( prob=p , size=1 ) ,
    
    logit(p) <- 
        a +
        a_survey[country_year_id] +  
        b_slept * slept ,  
    
    a ~ dnorm(0, 10),
    a_survey[country_year_id] ~ dnorm( 0 , 10 ),
    b_slept ~ dnorm(0, 10) 
)

fit.m.country = map( m.country , data = d, 
                     start = list( a = log(.3) , a_survey = log(1),
                                   b_slept = log(1) ) )

save(d,fit.m.country, file = "fit.m.country.rda")
load("fit.m.country.rda")

exp( precis( fit.m.country, depth = 2 )@output )

#@ log
m.country.log <- alist(
    hml3235 ~ dbinom(  size=1 , prob=p ) ,
    
    log(p) <- 
        a +
        a_survey[country_year_id] +  
        b_slept * slept ,  
    
    a ~ dnorm(0, 1),
    a_survey[country_year_id] ~ dnorm( 0 , 10 ),
    b_slept ~ dnorm(0, 1) 
)

fit.m.country.log = map2stan( m.country.log , data = d, iter = 200)
# fit.m.country.log = map( m.country.log , data = d  , 
#                      start = list( a = log(.3) , a_survey = log(1),
#                                    b_slept = log(1) ) )
plot(fit.m.country.log)

save(d,fit.m.country.log, file = "fit.m.country.log.rda")
load("fit.m.country.log.rda")

exp( precis( fit.m.country.log, depth = 2 )@output )

# refit with fewer iterations...
fit.m.country.log.short = map2stan( m.country.log , data = d, iter = 200)

# model fitting and convergence:
# compare(fit.m.country.log, ...)
post = extract.samples(fit.m.country.log )
pred = predict(fit.m.country.log )

x = rep( 0:1, 1000)
y = post$a + x*post$b_slept
plot(0,0, type = "n", xlab="slept under net", ylab = "parasitemia", ylim = 0:1,
     xaxt = "n", xlim = 1:2
)
axis( 1, at = 1:2, labels = c("did not", "did") )
lines(a:2,)

```







## Multi-level, median a postiori rethinking package

estimates both risk (intercept)  and risk ratio--in this case OR because using binomial link function.  

```{r}


  #calc year-month from CMC date
Year = 1900 + floor( (x$hv008 - 1)/ 12 )
Month = x$hv008 - 12 * (Year - 1900)

country = coerce_index( x$iso3c )

urban_rural = x$hv025

psu = paste0( x$iso3c, Year, x$hv001  ) 
psu_id = coerce_index( psu )

household = paste0( x$iso3c, Year, x$hv001, x$hv002 ) 
household_id = coerce_index( household )

# ML0 fit country
d = as.data.frame( cbind( hml3235, slept, country) )
ml0 <- alist(
    hml3235 ~ dbinom( prob=p , size=1 ) ,
    logit(p) <- 
        a_country[country] +  
        b_slept * slept ,  
    a_country[country] ~ dnorm( 0 , 10 ),
    b_slept ~ dnorm(0, 10) 
)
fitml0 = map( ml0 , data = d)
exp( precis(fitml0)@output )

# fit country and urban/rural
d = as.data.frame( cbind( hml3235, slept, country, urban_rural ) )

ml0.1 <- alist(
    hml3235 ~ dbinom( prob=p , size=1 ) ,
    logit(p) <- a + 
        a_country[country] +  
        a_urban[urban_rural] +
        b_slept * slept ,  
    a_country[country] ~ dnorm( 0 , 10 ),
    a_urban[urban_rural] ~ dnorm( 0 , 10 ),
    c(a, b_slept) ~ dnorm(0, 10) 
)
fitml0.1 = map( ml0.1 , data = d)
exp( precis(fitml0.1)@output )

compare(fitml0, fitml0.1)

# Map2Stan  ####
d = as.data.frame( cbind( hml3235, slept, country) )
ml1 <- alist(
    hml3235 ~ dbinom(1, p) ,
    logit(p) <- a + 
        a_country[country] +  
        # a_psu[psu_id] +  
        # a_household[household_id] + 
        b_slept * slept ,  
    
    a_country[country] ~ dnorm( 0 , sigma_country ),
    # a_psu[psu_id] ~ dnorm( 0 , sigma_psu ),
    # a_household[household_id] ~ dnorm( 0 , sigma_household ),
    
    c(a, b_slept) ~ dnorm(0, 10) ,
    
    sigma_country ~ dcauchy(0,1)
    # sigma_psu ~ dcauchy(0,1)
    # sigma_household ~ dcauchy(0,1)
)

# first fit 
fit_ml1 = map2stan( ml1 , data = d, chains = 1, cores = 1) 
exp( precis(fit_ml1, depth = 1 )@output )

# full fit 
fit_ml1.full  = map2stan( fit_ml1 , data = d , chains = 4, cores = 2) 
exp( precis(fit_ml1.full, depth = 1 )@output)

saveRDS(fit_ml1.full , "fit_ml1.full.rds")

# m2 <- map2stan(  # TAKES 10 minutes to run!
#     alist(
#         hml3235 ~ dbinom(1, theta),
#         logit(theta) = a + bs * slept ,
#         a ~ dnorm(0,10),
#         bs ~ dnorm(0,10)
#     ) ,
#     data=d, chains=2, cores=1 )
# 
# exp( precis(m2)@output )

# library(arm)
# lmer( hml3235 ~ slept , data = d)

```


## Visualize parasitemia versus bednet coverage by cluster

Because the samples are paired, need to be sure that each part of the pair exists.  For a given level on x-axis, there may be only groups from high prevalence areas that all use nets, so there is no no-net group, that get compared with groups from low-prevalence areas where there are none in the use-net group.  So, the loess line is comparing groups using nets in high prevalence areas to groups not using nets in low prevalence areas; both groups may have roughly same level of risk, therefore it looks as if the nets are not protective.

Possible solutions:  
1. make sure that each cluster has a parasitemia rate for no-net and net group (filter on complete.cases() )


## Group by cluster (hv001)

```{r}
 d1  = x %>% 
  mutate( 
    
    parasitemia = ifelse( hml32 %in% 1 | hml35 %in% 1, 1, ifelse( hml32 %in% 0 | hml35 %in% 0, 0, NA) ),
    
    know_parasitemia = parasitemia %in% 0:1 ,
    
    know_net_use =  hml19 %in% 0:1 | hml20 %in% 0:1 ,
    
    used_net = hml19 %in% 1 | hml20 %in% 1
    
          ) %>%
  
  group_by( country, survey, year, hv001) %>%
  
  summarise(
    
    pfpr = sum( parasitemia, na.rm = TRUE ) / sum( know_parasitemia, na.rm = TRUE ),
    
    slept_under_net = sum( used_net, na.rm = TRUE ) / sum( know_net_use, na.rm = TRUE ),
    
    # net_per_person = sum( hml1, na.rm  = TRUE ) / sum( hv012, na.rm  = TRUE ),
    
    pfpr_net = sum( parasitemia & used_net, na.rm = TRUE ) / sum( know_parasitemia & used_net, na.rm = TRUE ),
    
    pfpr_no_net = sum( parasitemia & !used_net, na.rm = TRUE ) / sum( know_parasitemia & !used_net, na.rm = TRUE ),
    
    n = n(),
    n_para = sum( know_parasitemia, na.rm = TRUE ),
    n_net_para = sum( know_parasitemia & used_net, na.rm = TRUE ) ,
    n_no_net = sum( know_parasitemia & !used_net, na.rm = TRUE )
  ) 

```

```{r}
ggplot( d1, aes( x = slept_under_net, y = pfpr)) +
  geom_point() +
  geom_smooth()

```

# Visualize parasitemia versus bednet coverage as Number of Nets per person 

```{r, eval=FALSE}

ggplot( d1 %>% filter( net_per_person<=1 ), aes( x = net_per_person, y = pfpr)) +
  geom_point() +
  geom_smooth()
```

##  Relative risk for sleeping under net

```{r}
gather_cols = c( "pfpr_net" , "pfpr_no_net")

d1. = d1 %>% ungroup %>% 
  filter( complete.cases(pfpr_net , pfpr_no_net)) %>%
  filter( net_per_person<=1 ) %>%
  gather_("cohort", "val" , gather_cols ) %>%
  arrange( country, survey, year, hv001 )

```

```{r}

ggplot( d1., aes( x = slept_under_net, y = val, color = cohort, group = cohort)) +
  geom_point() +
  geom_smooth()
```

```{r}

ggplot( d1. , aes( x = net_per_person, y = val, color = cohort, group = cohort)) +
  geom_point() +
  geom_smooth()
```

# Same analysis by HOUSEHOLD (hv002), rather than PSU

```{r}
 d2 = x %>% 
  mutate( 
    
    parasitemia = ifelse( hml32 %in% 1 | hml35 %in% 1, 1, ifelse( hml32 %in% 0 | hml35 %in% 0, 0, NA) ),
    
    know_parasitemia = parasitemia %in% 0:1 ,
    
    know_net_use =  hml19 %in% 0:1 | hml20 %in% 0:1 ,
    
    used_net = hml19 %in% 1 | hml20 %in% 1
    
          ) %>%
  
  group_by( country, survey, year, hv001, hv002) %>%
  
  summarise(
    
    pfpr = sum( parasitemia, na.rm = TRUE ) / sum( know_parasitemia, na.rm = TRUE ),
    
    slept_under_net = sum( used_net, na.rm = TRUE ) / sum( know_net_use, na.rm = TRUE ),
    
    net_per_person = sum( hml1, na.rm  = TRUE ) / sum( hv012, na.rm  = TRUE ),
    
    pfpr_net = sum( parasitemia & used_net, na.rm = TRUE ) / sum( know_parasitemia & used_net, na.rm = TRUE ),
    
    pfpr_no_net = sum( parasitemia & !used_net, na.rm = TRUE ) / sum( know_parasitemia & !used_net, na.rm = TRUE ),
    
    n = n(),
    n_para = sum( know_parasitemia, na.rm = TRUE ),
    n_net = sum( know_net_use, na.rm = TRUE ),
    n_net_para = sum( know_parasitemia & used_net, na.rm = TRUE ) ,
    n_no_net = sum( know_parasitemia & !used_net, na.rm = TRUE )
  ) %>%
  
  mutate( rr_slept_under_net =  pfpr / slept_under_net )


```


```{r}
ggplot( d2, aes( x = slept_under_net, y = pfpr)) +
  geom_point() +
  geom_smooth() 

```

```{r}

ggplot( d2 %>% filter( net_per_person<=1 ), aes( x = net_per_person, y = pfpr)) +
  geom_point() +
  geom_smooth() 
```

## relative risk of sleeping under net by percent slept under net

```{r}
ggplot( d2 %>% filter(slept_under_net>0), 
        aes( x = slept_under_net, 
             y = rr_slept_under_net)
        ) +
  geom_point() +
  geom_smooth() 

```

## relative risk of sleeping under net by number of nets / person

```{r}
ggplot( d2 %>% filter(slept_under_net>0 , 
                      net_per_person<=1 ), 
        aes( x = net_per_person, 
             y = rr_slept_under_net)) +
  geom_point() +
  geom_smooth() 

```

## cohorts

```{r}
gather_cols = c( "pfpr_net" , "pfpr_no_net")

d2. = d2 %>% ungroup %>% 
  filter( complete.cases(pfpr_net , pfpr_no_net)) %>%
  filter( net_per_person<=1 ) %>%
  gather_("cohort", "val" , gather_cols ) %>%
  arrange( country, survey, year, hv001, hv002 )

```

## Percent slept under net
```{r}
ggplot( d2., aes( x = slept_under_net, y = val, color = cohort, group = cohort)) +
  geom_point() +
  geom_smooth( ) + 
  scale_y_continuous("PfPR", limits = c(0,1)) 
```

## Number net per person
```{r}

ggplot( d2. , aes( x = net_per_person, y = val, color = cohort, group = cohort)) +
  geom_point() +
  scale_y_continuous("PfPR", limits = c(0,1))  +
  geom_smooth() 
```


## by country
```{r, warning=FALSE, message= FALSE}

ggplot( d2. , aes( x = net_per_person, y = val, color = cohort, group = cohort)) +
  geom_point() +
  geom_smooth() +
  scale_y_continuous("PfPR", limits = c(0,1))  +
  facet_wrap( ~country)
```

## by prevalence
```{r, warning=FALSE, message=FALSE}

ggplot( d2. %>%
  mutate(
        prevalence = cut( pfpr, 3) 
  )
, aes( x = slept_under_net, y = val, color = cohort, group = cohort)) +
  geom_point(alpha = .5) +
  geom_smooth() +
  scale_y_continuous("Parasitemia", limits = c(0,1)) +
  facet_wrap( ~ prevalence)
```


```{r, warning=FALSE, message=FALSE}

ggplot( d2. %>%
  mutate(
        prevalence = cut( pfpr, 3) 
  )
  , aes( x = net_per_person, y = val, color = cohort, group = cohort)) +
  geom_point(alpha = .5) +
  geom_smooth() +
  scale_y_continuous("Parasitemia", limits = c(0,1)) +
  facet_wrap( ~ prevalence)
```

## by Year

```{r, warning=FALSE, message=FALSE}

ggplot( d2. %>%
  mutate(
        YRs = cut( as.numeric(substr(year, 1, 4)) , breaks = c(2006, 2010, 2012, 2015 ) )
  ) %>% filter ( !is.na(YRs) )
, aes( x = slept_under_net, y = val, color = cohort, group = cohort)) +
  geom_point(alpha = .5) +
  geom_smooth() +
  scale_y_continuous("Parasitemia", limits = c(0,1)) +
  facet_wrap( ~ YRs)
```


```{r, warning=FALSE, message=FALSE}

ggplot( d2. %>%
  mutate(
        YRs = cut( as.numeric(substr(year, 1, 4)) , breaks = c(2006, 2010, 2012, 2015 ) )
  ) %>% filter ( !is.na(YRs) )
  , aes( x = net_per_person, y = val, color = cohort, group = cohort)) +
  geom_point(alpha = .5) +
  geom_smooth() +
  scale_y_continuous("Parasitemia", limits = c(0,1)) +
  facet_wrap( ~ YRs)
```

## by YRs and prevalence

```{r, warning= FALSE}
ggplot( d2. %>%
  mutate(
        YRs = cut( as.numeric(substr(year, 1, 4)) , breaks = c(2006, 2010, 2012, 2015 ) ),
        prevalence = cut( pfpr, 3) 
  ) %>% filter ( !is.na(YRs) )
  , aes( x = slept_under_net, y = val, color = cohort, group = cohort)) +
  geom_point(alpha = .5) +
  geom_smooth() +
  scale_y_continuous("Parasitemia", limits = c(0,1)) +
  facet_wrap( prevalence ~ YRs)
```
 
```{r, warning= FALSE}
ggplot( d2. %>%
  mutate(
        YRs = cut( as.numeric(substr(year, 1, 4)) , breaks = c(2006, 2010, 2012, 2015 ) ),
        prevalence = cut( pfpr, 3) 
  ) %>% filter ( !is.na(YRs) )
  , aes( x = net_per_person, y = val, color = cohort, group = cohort)) +
  geom_point(alpha = .5) +
  geom_smooth() +
  scale_y_continuous("Parasitemia", limits = c(0,1)) +
  facet_wrap( prevalence ~ YRs)
```
